---
title: "Data Viz Exploration"
author: 
- "Gabriel Krotkov"
- 'Style Guide: Modified Tidyverse Style Guide'
date: "Due Thursday, Sept 28th, 2023 (11:59 PM EDT) on Gradescope"
output:
  pdf_document:
    toc: yes
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

My variant of the Tidyverse style guide can be found [here](https://github.com/GKrotkov/gkrotkov_style_r/tree/main)

```{r}
library(tidyverse)
library(hexbin)
```

# Data Loading

```{r}
# prepend the string prefix onto each of the items of the vector v, 
# unless the item already starts with prefix
prepend_unless <- function(v, prefix, sep = "_"){
    idx <- !startsWith(v, prefix)
    v[idx] <- paste(prefix, v[idx], sep = sep)
    return(v)
}

linkedin <- read_csv("archive/job_postings.csv", 
                     show_col_types = FALSE)
companies <- read_csv("archive/company_details/companies.csv", 
                     show_col_types = FALSE)
colnames(companies) <- prepend_unless(colnames(companies), "company")
industries <- read_csv("archive/company_details/company_industries.csv", 
                     show_col_types = FALSE)
employee_counts <- read_csv("archive/company_details/employee_counts.csv", 
                     show_col_types = FALSE)

linkedin <- merge(linkedin, companies, by = "company_id")
linkedin <- merge(linkedin, industries, by = "company_id")

top_ten <- table(linkedin$industry)[order(table(linkedin$industry), 
                                          decreasing = TRUE)][1:10]
```

```{r}
f500 <- read_tsv("fortune_500.txt", col_names = FALSE)
colnames(f500) <- "company_name"
f500 <- f500$company_name
```

we want to create a new dataframe with "company" as the rows:
- need to be able to see pct of employees who are full time. 
- avg, median, sd salary
- employee headcount over time
- headquarters country
- company size

# Script to generate the target dataframe

todo list for linkedin data: 
- formatted experience level

Decided to discard some information from work_type - we don't care how many volunteer and internship roles they put up.

How do I summarise a categorical variable? (like pay period, work type, etc.)

```{r}
# prepend the string prefix onto each of the items of the vector v, 
# unless the item already starts with prefix
prepend_unless <- function(v, prefix, sep = "_"){
    idx <- !startsWith(v, prefix)
    v[idx] <- paste(prefix, v[idx], sep = sep)
    return(v)
}

linkedin <- read_csv("archive/job_postings.csv", 
                     show_col_types = FALSE)
cidx <- which(colnames(linkedin) == "formatted_experience_level")
colnames(linkedin)[cidx] <- "experience"
# compress linkedin data into company-level information
linkedin <- linkedin %>%
    group_by(company_id) %>%
    summarise(n_postings = n(), 
              mean_salary_max = round(mean(max_salary, na.rm = TRUE), 2), 
              mean_salary_med = round(mean(med_salary, na.rm = TRUE), 2), 
              mean_salary_min = round(mean(min_salary, na.rm = TRUE), 2), 
              yearly_pct = round(sum(pay_period == "YEARLY") / n(), 2), 
              monthly_pct = round(sum(pay_period == "MONTHLY") / n(), 2), 
              hourly_pct = round(sum(pay_period == "HOURLY") / n(), 2), 
              fulltime_pct = round(sum(work_type == "FULL_TIME") / n(), 2), 
              parttime_pct = round(sum(work_type == "PART_TIME") / n(), 2), 
              entry_pct = round(sum(experience == "Entry Level") / n(), 2),
              assoc_pct = round(sum(experience == "Associate") / n(), 2),
              mid_pct = round(sum(experience == "Mid-Senior level") / n(), 2), 
              director_pct = round(sum(experience == "Director") / n(), 2),
              exec_pct = round(sum(experience == "Executive") / n(), 2),
              intern_pct = round(sum(experience == "Internship") / n(), 2))

companies <- read_csv("archive/company_details/companies.csv", 
                     show_col_types = FALSE)
colnames(companies) <- prepend_unless(colnames(companies), "company")

industries <- read_csv("archive/company_details/company_industries.csv", 
                     show_col_types = FALSE)
industries <- distinct(industries)

employee_counts <- read_csv("archive/company_details/employee_counts.csv", 
                     show_col_types = FALSE)
# compress 'employee_counts' into company on the rows
employee_counts <- employee_counts %>%
    group_by(company_id) %>%
    summarise(employees = mean(employee_count), 
              followers = mean(follower_count))

# attach compressed dataframes to "companies"
companies <- merge(companies, employee_counts, by = "company_id")
companies <- merge(companies, industries, by = "company_id")
companies <- merge(companies, linkedin, by = "company_id")

save(companies, "data.rda")
```

